---
title: "Tuesday LiDAR Lesson"
author: "Enrique Monta&ntilde;o"
date: "June 21, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Classification of Raster from Thresholds

First load the libraries

### NOTE

* require provides an error if the package is not installed

```{r load-libraries}
#load needed libraries
library(raster)
library(rhdf5)
library(rgdal)


```

## Load the Canopy Height Model Raster

Load the Canopy Height Model, simply the residual between the canopy structure and terrain

```{r import-chm-raster}
#load the canopy height model
chm <- raster("../NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarCHM.tif")

```

## Visualize the raster

Values of 0 on a landscape indicate regions without vegetation.  We need to set those pixels to `NA` for the sake of visualization.

```{r plot-chm}
#first set no data values
chm[chm == 0] <- NA

# plot the raster
plot(chm,
     main = "First Plot of TEAK CHM")

```


## Look at a Histogram of the values

Look at the distribution of canopy height values

```{r plot-hist}
#plot a histogram of chm
hist(chm,
     main="Distribution of Canopy Height  \n Lower Teakettle, California",
     xlab="Tree Height (m)",
     col="springgreen")

```

## Import Aspect

```{r import-aspect}

# import an aspect raster
aspect <- raster("../NEONdata/D17-California/TEAK/2013/lidar/TEAK_lidarAspect.tif")

# plot the aspect
plot(aspect,
main = "Aspect data for TEAK Site")

```


##  Create a Classification Matrix

* North Facing Slopes: 0-45 degrees, 315-360 degrees; class=1

* South Facing Slopes: 135-225 degrees; class=2

```{r class-matrix}
# create a vector of range and classes
class_m <- c(0, 45, 1, 
             45, 135, NA,
             135, 225, 2,
             225, 315, NA,
             315, 360, 1)

# make it a matrix
rcl_m <- as.matrix(class_m,
                   ncol = 3,
                   byrow = TRUE)
rcl_m
```

## Reclassify the Aspect

```{r reclass-raster}

# reclassify the aspect
asp_ns <- reclassify(aspect, rcl_m)

# plot the classified raster
plot(asp_ns,
     col=c("white","blue","green"), # hard code colors, unclassified (0)=white,
		 #N (1) =blue, S(2)=green
     main="North and South Facing Slopes \nLower Teakettle",
     legend=F)
# allow legend to plot outside of bounds
par(xpd=TRUE)
# create the legend
legend((par()$usr[2] + 20), 4103300,  # set x,y legend location
       legend = c("North", "South"),  # make sure the order matches the colors, next
       fill = c("blue", "green"),
       bty="n") # turn off border
```

## Save the output

```{r save-raster, eval=FALSE}

# save the output as geotiff
writeRaster(asp_ns,
            file = "../outputs/TEAK/TEAK_nsAspect2.tif",
            format="GTiff", #format
            options="COMPRESS=LZW", #compress the data
            overwrite = TRUE, #overwrite if exists
            NAflag = -9999) # standard no data value

```

## Mask the data

We often want to exclude data we aren't interested in for analysis.  A mask is a simple math operation multiplying a 1 (for values we want to keep) by the data raster.

```{r mask-raster}
# make a mask from NDVI
# import ndvi raster
ndvi <- raster("../NEONdata/D17-California/TEAK/2013/spectrometer/veg_index/TEAK_NDVI.tif")

# plot the ndvi
plot(ndvi,
     main = "NDVI for TEAK Site")

# Mask the NDVI using N/S slopes
nFacing_ndvi <- mask(ndvi, 
                     asp_ns)

plot(nFacing_ndvi, 
     main = "Masked NDVI")
```



